---
title: "Question 3"
author: "Aaron Quach"
date: "March 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('dplyr')
library('knitr')
library('stringr')
library('ggplot2')
source("Question2.R")
```

## Question 3 Analysis

### Which departments at each of these schools offer classes that have relatively high class GPAs and low GPAs? 

```{r uw_high_avg, echo = FALSE}
uw_high_avg

```
```{r vt_high_avg, echo = FALSE}
vt_high_avg

```
```{r uw_low_avg, echo = FALSE}
uw_low_avg

```
```{r vt_low_avg, echo = FALSE}
vt_low_avg

```

The first graph displays the top 6 departments at the at the University of Washington 
where  a number of classes in each of these departments have an average GPA at or above a 3.8. 
From this graph the English department and Sociology departments have the most number
of classes with 20 or more classes where students got high averages.

The second graph displays the top 6 departments at Virginia Tech where the
number of classes in each of these departments have an average GPA at or above a 3.8 and while 
none of the departments have more than 20 classes where students get high averages, 
the distribution is less as more than all 6 of the departments shown have more than
10 classes where the average GPA was or above a 3.8. However, for Virginia Tech, their 
top departments are Education Curriculum and Instruction and English. 

The third graph shows the top 6 departments at the University of Washington where
the number of classes in each of these departments have an average GPA at or below a 2.8. 
In this graph, Math has the most amount of classes at 14 while the second highest is 
the Biology department which has 8 classes with a low average GPA. 

The fourth graph shows the top 6 departments at Virgina Tech where the
number of classes in each of these departments have an average GPa at or below a 2.8.
In this graph the Math department and Electrical Computer Engineer department 
have the most amount of classes which has similarities to the UW data as the
Math department also had the highest number of classes with low GPAs. 


